---
title: "Handson_Ex03"
author: "Hou Tao"
date: "3 Feb 2023"
date-modified: "`r Sys.Date()`"
execute: 
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

# Visualising Uncertainty

Visualizing the uncertainty of point estimates

## 1 Load Library and Import Data

```{r}
pacman::p_load(tidyverse, plotly, DT, ggdist, gganimate, crosstalk, ggdist)
```

```{r}
exam <- read_csv("data/Exam_data.csv", show_col_types = FALSE)
```

### 1.1 Visualizing the uncertainty of point estimates: ggplot2 methods

-   group the observation by RACE,
-   computes the count of observations, mean, standard deviation and standard error of Maths by RACE, and
-   save the output as a tibble data table called my_sum.

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n = n(),
    mean = mean(MATHS),
    sd = sd(MATHS)
    ) %>%
  mutate(se = sd / sqrt(n - 1))

knitr::kable(head(my_sum), format = 'html')

```

### 1.2 Visualizing the uncertainty of point estimates: ggplot2 methods

to reveal the standard error of mean maths score by race.

```{r}
my_sum %>%
    ggplot(aes(x = RACE, y = mean)) + 
      geom_errorbar(
        aes(x = RACE,
            ymin = mean - se,
            ymax = mean + se),
        width = 0.2,
        colour = "black",
        alpha = 0.95,
        linewidth = 0.5) +
      geom_point(aes(x = RACE, y = mean),
                 stat = "identity",
                 color = "red",
                 size = 1.5,
                 alpha = 1) +
      ggtitle("Standard error of mean maths score by race")
  
```

### 1.3 Visualizing the uncertainty of point estimates: ggplot2 methods

```{r}

my_sum %>%
  mutate(RACE = fct_reorder(RACE, desc(mean))) %>%
  ggplot() +
    geom_errorbar(
      aes(x = RACE,
          ymin = mean - se * 1.96,
          ymax = mean + se * 1.96),
      width = 0.2,
      colour = "black",
      alpha = 0.9,
      linewidth = 0.5) +
    geom_point(aes(x = RACE, y = mean),
               stat = "identity",
               color = "red",
               size = 1.5,
               alpha = 1) +
    ggtitle("Standard error of mean maths score by rac")

```

### 1.4 Visualizing the uncertainty of point estimates with interactive error bars

```{r}

d <- highlight_key(my_sum)

t <- DT::datatable(d)

p <- my_sum %>%
      mutate(RACE = fct_reorder(RACE, desc(mean))) %>%
      ggplot() +
        geom_errorbar(
          aes(x = RACE,
              ymin = mean - se * 2.576,
              ymax = mean + se * 2.576),
          width = 0.2,
          colour = "black",
          alpha = 0.9,
          linewidth = 0.5) +
        geom_point(aes(x = RACE, y = mean),
                   stat = "identity",
                   color = "red",
                   size = 1.5,
                   alpha = 1) +
        ggtitle("Standard error of mean maths score by rac")

g <- highlight(ggplotly(p), off="plotly_doubleclick")

bscols(widths = c(5, 7), g, t)


```

# 2. Visualising Uncertainty: ggdist package

### 2.1 Visualizing the uncertainty of point estimates: ggdist methods

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) + 
  stat_pointinterval(.width = 0.95) + 
  labs(
    title = "Visiualising confidence internvals of mean math score",
    subtitle = "Mean Point + Multiple-Interval plot"
  )


```

### 2.2 Visualizing the uncertainty of point estimates: ggdist methods

by showing 95% and 99% confidence intervals.

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) + 
  stat_pointinterval(.width = c(0.95, 0.99)) + 
  labs(
    title = "Visiualising confidence internvals of mean math score",
    subtitle = "Mean Point + Multiple-Interval plot"
  )


```

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) + 
  stat_pointinterval(point_interval = median_qi) + 
  labs(
    title = "Visiualising confidence internvals of mean math score",
    subtitle = "Mean Point + Multiple-Interval plot"
  )
```

### 2.3 Visualizing the uncertainty of point estimates: ggdist methods

```{r, warning=FALSE}

exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_gradientinterval(
    fill = "skyblue",
    show.legend = TRUE
  ) + 
  labs(
    title = "Visualising confidence internvals of mean math score",
    substile = "Gradient + interval plot"
  )

```

### 2.4 Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)

```{r}
pacman::p_load(ungeviz)
```

```{r, warning=FALSE}
## devtools::install_github("wilkelab/ungeviz")


ggplot(data = exam,
       (aes(x = factor(RACE), y = MATHS))) +
  geom_point(
    position = position_jitter(height = 0.3, width = 0.05), 
    size = 0.4,
    color = "#0072B2",
    alpha = 1/2) + 
  geom_hpline(data = sampler(25, group = RACE), height = 0.6, color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)


```

### 2.5 Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)

```{r, warning=FALSE}

ggplot(data = exam,
       (aes(x = factor(RACE), y = MATHS))) +
  geom_point(
    position = position_jitter(height = 0.3, width = 0.05), 
    size = 0.4,
    color = "#0072B2",
    alpha = 1/2) + 
  geom_hpline(data = sampler(25, group = RACE), height = 0.6, color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)
```
