---
title: "Hands-on Exercise 2"
author: "Hou Tao"
---

# Creating elegent graphics with ggplot2

## Beyond ggplot2

installing and loading the required libraries

```{r}
pacman::p_load(tidyverse, patchwork, ggthemes, hrbrthemes, ggrepel, states, extrafont)
```

import data

```{r}
exam_data <- read_csv("data/Exam_data.csv", show_col_types = FALSE)
```

beyond ggplot2 annotation

```{r}
ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point() + 
  geom_smooth(formula = y ~ x, method = lm, size = 0.5) + 
  geom_label(aes(label = ID), hjust = 0.5, vjust = 0.5) + 
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  theme_economist() +
  ggtitle("English scores versus Maths scores for Primary 3")
```

ggplot2 annotation with *repel*, *ggrepel* provides *geoms* for ggplot2 to repel overlapping text labels, Text labels repel away from each other, away from data points, and away from edges of the plotting area (panel). - geom_text_repel() - geom_label_repel()

```{r}
ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point() + 
  geom_smooth(formula = y ~ x, method = lm, size = 0.5) + 
  geom_label_repel(aes(label = ID), fontface = "bold", max.overlaps = 30) + 
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  theme_economist() +
  ggtitle("English scores versus Maths scores for Primary 3")
```

Beyond ggplot2 themes

```{r}
ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  theme_economist() +
  ggtitle("Distribution of Maths scores")
```

Working with hrbrthemes package

hrbrthemes package provides a base theme that focus on typograhic elements.

```{r}

extrafont::loadfonts(device="win")

windowsFonts(Times=windowsFont("TT Times New Roman"))

ggplot(data = exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(base_family = "Times New Roman") 
```

Byond ggplot2 facet

```{r}

p1 <- ggplot(data = exam_data, aes(x = MATHS)) +
        geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
          coord_cartesian(xlim=c(0, 100)) +
            ggtitle("Distribution of Maths scores")


p2 <- ggplot(data = exam_data, aes(x = ENGLISH)) +
        geom_histogram(bins = 20, boundary = 100, color = "grey25", fill = "grey90") +
          coord_cartesian(xlim=c(0, 100)) +
            ggtitle("Distribution of English scores")


p3 <- ggplot(data = exam_data, aes(x = MATHS, y = ENGLISH)) +
        geom_point(color="grey25", fill="grey90") + 
          geom_smooth(formula = y ~ x, method = lm, size = 0.5) + 
              coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
                ggtitle("English scores versus Maths scores for Primary 3")

```

```{r}
p1 + p2 / p3
```

```{r}
(p1 / p2) | p3
```

```{r}
((p1 / p2) | p3) +
  plot_annotation(tag_levels = 'I')
```

```{r}
patchwork <- ((p1 / p2) | p3) + plot_annotation(tag_levels = 'I')
patchwork & theme_economist()

```

With inset_element() of patchwork, we can place one or several plots or graphic elements freely on top or below another plot.

```{r}
p3 + inset_element(p2, left = 0.02, bottom = 0.7, right = 0.5, top = 1)
```
